Parsing Postfix log files

### 2008-04-08 doc/logparser.tex
Improved the emergent behaviour section (I hope) and added a piece to the conclusion about the unusual flow of control.

XXX EITHER PICK A BETTER WORD OR GIVE A BETTER EXPLANATION OF EMERGENT
BEHAVIOUR --- SECTIONS 5.4.4 AND 8\@.

@@

The rules and actions exhibit emergent behaviour~\cite{Wikipedia-Emergence}
which is far more complicated than the behaviour encoded in the algorithm.
The top-level parser could be written in pseudo-code like so (with
indentation level indicating flow control):

@@

The solution to the complication in \sectionref{out of order log lines}
(out of order log lines) requires a list of the acceptable combinations of
Postfix programs a mail can pass through; however this does not correspond
to the \textit{actions\/} a mail must pass through.  Before that
complication was overcome, i.e.\ for the majority of the development of
this parser, there was nothing in the parser explicitly encoding the states
or paths a mail could take; no other component in the parser has any need
for this flow of data to be specified.  This eases the process of adding
new actions (described in \sectionref{adding new actions}), as there is no
requirement to explicitly insert the action into an algorithmic version of
the flow chart.

@@

XXX EXTEND THIS\@; SOMETHING ABOUT THE CONTROL FLOW BEING UNUSUAL, I.E.\
RULE DETERMINES ACTION, RATHER THAN ACTION EXPECTING INPUT\@.

The emergent behaviour~\cite{Wikipedia-Emergence} (\sectionref{Emergent
behaviour}) exhibited by the rules and actions is also interesting, and is
discussed after the flow chart (\sectionref{flow-chart}) and explanation of
the paths through the parser.  This emergent behaviour greatly eases adding
new actions (\sectionref{adding new actions}), as the actions do not have
to be inserted into an explicit flow of control; new actions will naturally
find their place in the paths through the parser available to a mail.

>>>>

@@

The rules and actions exhibit emergent behaviour~\cite{Wikipedia-Emergence}
which is far more complicated than the behaviour explicitly encoded in the
algorithm.  The top-level parser could be written in pseudo-code like so
(with indentation level indicating flow control):

@@

The actions introduce some of the complexity seen in the flow chart: all
explicit branches and the transitions between some stages (19--20,
21--22--23--24) in the second half of the flow chart
(figure~\refwithpage{flow chart image part 2}) are encoded within the
actions; the transitions between the other stages (16--17, 24--25) in the
second half arise from the interaction between actions

However the transitions between stages in the first half of the flow chart
(figure~\refwithpage{flow chart image part 1}) are not encoded anywhere in
the parser; the transitions are determined by the ordering of log lines in
the log file.  The complexity of the first 12 stages in the flow chart
emerges from the interaction of simple rules and actions, easing the
process of adding new actions (described in \sectionref{adding new
actions}), as there is no requirement to explicitly insert the action into
an algorithmic version of the flow chart.\footnote{The solution to the
complication in \sectionref{out of order log lines} (out of order log
lines) requires a list of the acceptable combinations of Postfix programs a
mail can pass through; however this does not correspond to the
\textit{parser actions\/} a mail must pass through.  Before that
complication was overcome, i.e.\ for the first half of the development of
this parser, there was nothing in the parser explicitly encoding the states
or paths a mail could take; no other component in the parser has any need
for this flow of data to be specified.}

@@

The emergent behaviour~\cite{Wikipedia-Emergence} (\sectionref{Emergent
behaviour}) exhibited by the rules and actions is also interesting, and is
discussed after the flow chart (\sectionref{flow-chart}) and explanation of
the paths through the parser.  This emergent behaviour greatly eases the
process of adding new actions (\sectionref{adding new actions}), as the
actions do not have to be inserted into an explicit flow of control; new
actions will naturally find their place in the paths through the parser.

The flow of control in this parser is quite different from that of most
other parsers.  In most traditional parsers the parser has a current state,
and each state has a fixed set of acceptable next states, with unacceptable
states causing parsing to fail --- e.g.\ when a \textbf{C} parser sees the
keyword \textbf{for} it expects to immediately see a left parenthesis, with
any other input causing parsing to fail.\footnote{Comments will already
have been removed by the preprocessor.}  This parser is different: the
rule which matches the next log line for a mail dictates the action that
will be invoked, which is equivalent to the next state in other parsers;
thus the next log line for a mail dictates the next state for that mail.
